\documentclass[fleqn]{article} 
\usepackage{titlesec}
\usepackage{hyperref}
\usepackage{url}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{eqnarray}
\usepackage{multicol}
\usepackage{enumitem}
\usepackage{indentfirst}
%\usepackage[utf8]{inputenc}
%\usepackage[T1]{fontenc}
\setlength\columnsep{20pt}
\usepackage[margin={.75in,1.25in}]{geometry}

% Section spacing and font setup
\titlespacing{\section}{0pt}{10pt}{10pt}
\titleformat*{\section}{\large\bfseries}

%Bibliography setup
\usepackage[backend=biber, style=authoryear]{biblatex}
\addbibresource{references.bib}

\title{\huge Planetary Obliquity and Stellar Structure \\ \Large ASTR-154 Final Project}
\date{June 7, 2026}

\author{Babak Aryan\\Justin Kan}

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

%\nipsfinalcopy % Uncomment for camera-ready version

\begin{document}
	\maketitle

	\section*{ \centering \textbf{ABSTRACT}}
	
	Planetary obliquity, a planet’s axial tilt with respect to its plane of orbit, is the primary driver of seasons and climate cycles for an earth-like planet. An obliquity close to $0^\circ$, for example, eliminates seasonal changes; locations closer to the equator remain hot, while the poles stay perpetually frozen. A large obliquity ($> \; \sim\!45^\circ$), on the other hand, creates large variations in climate across different altitudes on the planet resulting in the hemispheres cycling between very hot summers and brutally cold winters over an orbital year.	According to (\cite{Vervoort}), one important condition to increase the probability of “the habitability of Earth-like planets” is for the mean obliquity to be “sufficiently low to maintain temperate temperatures over large parts of its surface throughout the orbital year.”  Planetary obliquity, therefore, should be considered as one of the essential parameters when searching for habitable exoplanets. Recent space telescope missions such as Kepler/K2, Tess and CoRoT have contributed to the discovery of $\sim40,000$ exoplanets cataloged in NASA Exoplanet Archive. Obliquity data, however, is available for only a handful ($274$) of these planets. Therefore, an efficient search for "habitable" exoplanets necessitates a better understanding of planetary obliquity population in order to better identify suitable candidates. Recent research on the evolution of planetary obliquity has shown the effect of non-single ("binary") stellar host systems on mean equilibrium value ($\epsilon_m$) as well as long term variation ($\Delta\epsilon$) of planetary obliquity (\cite{Huang}), (\cite{Quarles}).	In this paper, we explore the planetary obliquity populations as a function of the planet's host system's "binarity"\footnote{Throughout this paper "binarity" refers to the condition of a system containing more than 1 host star.}. We show that obliquity population for planets in single stellar host systems differ from the population for those in "binary" systems. In order to divide our exoplanet sample set into those in single star systems vs. those in binary systems, we rely on the number of listed stars in the data set as well as the test statistics RUWE (Re-normalized Unit Weight Error) which is a strong indicator of the binarity of a system (\cite{RUWE_Ginard}), (\cite{RUWE_Logan}). We show that, given a grouping of exoplanets based on both the number of stars listed and the RUWE number, yields obliquity populations which are distinctly different with a $p$ value of $< .03$. Given the high correlation between RUWE and binarity,
	

	\begin{multicols}{2}
		\section*{Motivation}
		What question(s) is your project attempting to answer, and why is this important to physics or astronomy? Give enough background that a nonspecialist can understand the motivation.
		
		\section*{Data}
		Describe the data set you used. Where did it come from and how was it obtained? What telescopes/instruments/software were used?
		
		\section*{Methods}
		Describe your analysis procedure. Why did you choose the techniques you used? If you used a technique beyond what we learned this quarter, teach it to the class. If there were other techniques you considered but chose not to use, explain why. 
		
		\section*{Results}
		What did you learn by applying your methods to your data? How does this help you answer the question(s) posed in the motivation? Include a discussion of the uncertainty or statistical significance of your results.
		
		\section*{Next Steps}
		If you were to continue along this line of research, how could you expand on your project to answer the question(s) better or to answer related questions? What data and techniques would you need? If your project didn't work out as expected, what do you think went wrong, and how could you improve?
	\end{multicols}
	
	\pagebreak
	
	\printbibliography

\end{document}

